---
title: "eda"
format: html
editor: visual
---

---
title: "Exploratory Data Analysis of U.S. Congress Legislators"
format: html
editor: visual
---

```{r setup, include=FALSE}
# Load necessary libraries
library(tidyverse)
library(lubridate)
library(jsonlite)
library(knitr)
library(kableExtra)
```


## Introduction

This report presents an exploratory data analysis (EDA) of the U.S. Congress legislators dataset, which includes information on members of Congress from 1789 to the present. The dataset is available in multiple formats (YAML, JSON, CSV) and provides details such as names, party affiliations, terms in office, and social media accounts.

## Data Loading

We will load the current legislators' data from the JSON file.

```{r}
# Load current legislators data
legislators_current <- fromJSON("https://unitedstates.github.io/congress-legislators/legislators-current.json", flatten = TRUE)

# Convert to tibble for easier handling
legislators_df <- as_tibble(legislators_current)
```

# Data Overview

Let's examine the structure and some sample entries of the dataset.

```{r}
# Display the structure of the dataset
str(legislators_df, max.level = 1)

# Show the first few rows
head(legislators_df, n = 5)
```

# Data Cleaning

We will extract relevant information such as full names, party affiliations, states, genders, and birthdays.

```{r}
# Extract relevant fields
legislators_clean <- legislators_df %>%
  mutate(
    full_name = paste(name.first, name.last),
    party = map_chr(terms, ~ .x[[length(.x)]]$party),
    state = map_chr(terms, ~ .x[[length(.x)]]$state),
    gender = bio.gender,
    birthday = ymd(bio.birthday),
    age = as.integer(interval(birthday, today()) / years(1))
  ) %>%
  select(full_name, party, state, gender, birthday, age)

```

